\chapter{Stochastic Processes}

\section{Definition and Basic Properties}

A stochastic process is a mathematical object defined to describe the evolution of a system subject to randomness over time. Formally, it is a collection of random variables defined on a common probability space.

\begin{definition}[Stochastic Process]
Let $(\Omega, \mathcal{F}, \mathbb{P})$ be a probability space. A \textbf{stochastic process} is a family of random variables $X = (X_t)_{t \in I}$ defined on $(\Omega, \mathcal{F})$ and taking values in a measurable space $(E, \mathcal{E})$, often called the \textit{state space}.
\end{definition}

The set $I$ is called the \textbf{index set} or \textbf{parameter set}, representing time.
\begin{itemize}
    \item \textbf{Discrete-time}: If $I$ is a countable set, such as $\mathbb{N} = \{0, 1, 2, \dots\}$ or a finite set $I = \{0, 1, \dots, T\}$. In this course, we primarily focus on the discrete-time setting with a finite horizon $I = \{0, 1, \dots, T\}$.
    \item \textbf{Continuous-time}: If $I$ is an interval of $\mathbb{R}$, such as $[0, T]$ or $[0, \infty)$. This is the subject of advanced stochastic calculus (e.g., Brownian motion).
\end{itemize}

The state space $E$ is typically $\mathbb{R}^d$ equipped with the Borel $\sigma$-algebra $\mathcal{B}(\mathbb{R}^d)$.

\section{The Canonical Construction}

One standard way to construct a stochastic process is to define it on a \textit{product space}. This is often referred to as the "canonical" construction, where the sample space $\Omega$ itself represents the set of all possible paths of the process.

Consider a finite horizon $T \in \mathbb{N}$ and a state space $E$.
\begin{itemize}
    \item Let $\Omega = E^T = \{ \omega = (\omega_1, \dots, \omega_T) : \omega_i \in E \}$ be the set of all possible sequences of length $T$.
    \item We can define a $\sigma$-algebra $\mathcal{F}$ on $\Omega$, typically the product $\sigma$-algebra.
    \item We define a probability measure $\mathbb{P}$ on $(\Omega, \mathcal{F})$.
\end{itemize}

In this setting, the stochastic process $X = (X_t)_{1 \leq t \leq T}$ can be defined as the \textbf{coordinate mapping}:
\[
X_t(\omega) = \omega_t, \quad \text{for } \omega = (\omega_1, \dots, \omega_T).
\]
Here, the randomness $\omega$ \textit{is} the entire trajectory of the process, and $X_t$ simply "reads" the value of the trajectory at time $t$.

\subsection{Example: Coin Toss (Random Walk)}

Consider the "Head or Tail" game where we toss a balanced coin $T$ times.
\begin{itemize}
    \item \textbf{State Space}: $E = \{H, T\}$ (Head, Tail). Alternatively, we can map this to numerical values, e.g., $\{1, 0\}$ or $\{1, -1\}$.
    \item \textbf{Sample Space}: $\Omega = E^T = \{ \omega = (\omega_1, \dots, \omega_T) : \omega_i \in \{H, T\} \}$. The size of the sample space is $|\Omega| = 2^T$.
    \item \textbf{Probability Measure}: Since the coin is balanced, every outcome sequence is equally likely.
    \[
    \mathbb{P}(\{\omega\}) = \frac{1}{|\Omega|} = \frac{1}{2^T}, \quad \forall \omega \in \Omega.
    \]
    The $\sigma$-algebra is the power set $\mathcal{F} = 2^\Omega$.
\end{itemize}

We can define random variables $X_t$ representing the outcome of the $t$-th toss:
\[
X_t(\omega) = \begin{cases} 1 & \text{if } \omega_t = H \\ 0 & \text{if } \omega_t = T \end{cases}
\]
The collection $(X_t)_{1 \leq t \leq T}$ is a stochastic process taking values in $\{0, 1\}$.

We can also define the \textbf{accumulated sum} process $S = (S_t)_{1 \leq t \leq T}$:
\[
S_t = \sum_{i=1}^t X_i
\]
The process $(S_t)$ takes values in the set $\{0, 1, \dots, T\}$ and represents the total number of "Heads" obtained up to time $t$.

\section{Trajectories (Sample Paths) and Notation}

The most complete and rigorous way to define a stochastic process is as a function of two variables: time and randomness.
\[
X : I \times \Omega \to E
\]
Let us break down this structural notation:
\begin{itemize}
    \item $X$: The name of the stochastic process.
    \item $I$: The \textbf{index set} (time domain), e.g., $\{0, 1, \dots, T\}$.
    \item $\Omega$: The \textbf{sample space} (the set of all possible random outcomes $\omega$).
    \item $\times$: The \textbf{Cartesian product}, meaning $X$ takes a pair of inputs $(t, \omega)$.
    \item $\to E$: The \textbf{state space}. The $\to$ arrow defines the structural mapping from the domain sets to the codomain set $E$.
\end{itemize}

When we write $X(t, \omega) = x$, we mean: "At time $t$, given that the random scenario $\omega$ occurred, the value of the process is $x \in E$."

This two-variable perspective allows us to "fix" one variable and study the process from two complementary views. To understand these views, it is crucial to distinguish between two types of arrows used in mathematical mapping:
\begin{itemize}
    \item The straight arrow \textbf{$\to$} (e.g., $A \to B$) defines the \textit{sets} involved: it maps a domain to a codomain.
    \item The maps-to arrow \textbf{$\mapsto$} (e.g., $x \mapsto f(x)$) defines the \textit{operation rule}: it shows exactly how a specific element from the domain is transformed into an element in the codomain.
\end{itemize}

\subsection{View 1: A collection of random variables (Fixing time)}
If we fix a specific time $t \in I$ and let $\omega$ vary, we obtain a new function:
\[
X(t, \cdot) : \Omega \to E, \quad \omega \mapsto X_t(\omega)
\]
Here, $X(t, \cdot)$ is a \textbf{Random Variable}. The dot "$\cdot$" indicates that the randomness variable is left free. We usually drop the dot and simply write \textbf{$X_t$}. It answers the question: "What will the value be at time $t$?" The answer is a random distribution over $E$.

\subsection{View 2: A random function / Trajectory (Fixing randomness)}
If the random outcome has already been realized (we fix a specific scenario $\omega \in \Omega$) and we let time run, we obtain:
\[
X(\cdot, \omega) : I \to E, \quad t \mapsto X_t(\omega)
\]
For a fixed outcome $\omega$, the deterministic function $t \mapsto X_t(\omega)$ is called a \textbf{trajectory}, \textbf{sample path}, or \textbf{realization} of the process.

\begin{definition}[Trajectory]
For each $\omega \in \Omega$, the function $X(\cdot, \omega)$ describes one specific history of the system's evolution over time.
\end{definition}

If we denote $\Sigma = E$ (the state space) and $\Sigma^I$ as the set of all functions from $I$ to $\Sigma$, then the entire process $X$ can be viewed as a single random variable taking values in the function space $\Sigma^I$:
\[
X: (\Omega, \mathcal{F}) \to (\Sigma^I, \mathcal{G}^{\otimes I})
\]
where $\mathcal{G}^{\otimes I}$ is the approproiate product $\sigma$-algebra on the function space.
